Beyond Diagonal RIS: MIMO Channel Shaping and Rate Maximization

### 2024-12-20 manuscript/main.tex
:construction: Updating structure

% \title{Channel Shaping Using Beyond Diagonal RIS:\\Analysis, Optimization, and Enhanced Flexibility}
% \title{MIMO Channel Shaping and Rate Maximization via Beyond Diagonal RIS}
\title{Beyond Diagonal RIS: MIMO Channel Shaping and Rate Maximization}
% \title{MIMO Channel Shaping Using Beyond Diagonal RIS: Singular Value Redistribution and Rate Maximization}
\author{
	\IEEEauthorblockN{
		Yang Zhao, \IEEEmembership{Member, IEEE,}
		Hongyu Li, \IEEEmembership{Graduate Student Member, IEEE,}\\
		Massimo Franceschetti, \IEEEmembership{Fellow, IEEE,}
		and Bruno Clerckx, \IEEEmembership{Fellow, IEEE}
	}
	\thanks{
		Yang Zhao, Hongyu Li, and Bruno Clerckx are with the Department of Electrical and Electronic Engineering, Imperial College London, London SW7 2AZ, U.K. (e-mail: \{\href{mailto:yang.zhao18@imperial.ac.uk}{yang.zhao18}, \href{mailto:c.li21@imperial.ac.uk}{c.li21}, \href{mailto:b.clerckx@imperial.ac.uk}{b.clerckx}\}@imperial.ac.uk).

@@

	% geodesic \gls{rcg} algorithm is tailored for smooth optimization problems of asymmetric \gls{bd}-\gls{ris} with arbitrary group size, then invoked for the Pareto frontier of channel singular values.
	% To understand the gain from off-diagonal entries, we also derive analytical singular value bounds in rank-deficient and fully-connected scenarios.
	% As a side product, we tackle \gls{mimo} rate maximization problem by alternating between active beamformer (eigenmode transmission) and passive beamformer (\gls{rcg} algorithm) until convergence.
	% A low-complexity suboptimal solution based on channel shaping is also proposed, where the decoupled problem is formulated as channel power maximization and solved in closed form iteratively.
	% % We then show how channel shaping decouples the design and amounts to a low-complexity suboptimal solution.
	% % A low-complexity suboptimal solution decoupling both blocks is also proposed, where the channel shaping subproblem is formulated as channel power maximization and solved in closed form iteratively.
	% Theoretical analysis and numerical evaluation reveal that the shaping advantage of \gls{bd}-\gls{ris} increases with group size and \gls{mimo} dimensions, stemming from stronger channel rearrangement and space alignment capabilities.
\end{abstract}

@@

		% First, we quantify the capability of a \gls{bd}-\gls{ris} to reshape the \gls{mimo} point-to-point channel in terms of singular values.
		% The \emph{Pareto frontiers} are characterized by optimizing the {weighted sum of singular values}, where the weights can be positive, zero, or negative.
		% % This generalizes most singular value metrics and provides a powerful design framework.
		% The resulting singular value region generalizes most relevant metrics and provides an intuitive channel shaping benchmark.
		% We then discuss some analytical singular value bounds in rank-deficient and fully-connected scenarios, which help to demystify the gain from off-diagonal entries.
		% This is the first paper to answer the channel shaping question and highlight the \gls{bd}-\gls{ris} gain from a Pareto perspective.

@@

	\emph{Notation:}
	Italic, bold lower-case, and bold upper-case letters indicate scalars, vectors and matrices, respectively.
	$\jmath$ denotes the imaginary unit.
	$\mathbb{C}$ represents the set of complex numbers.
	$\mathbb{H}^{n \times n}$, $\mathbb{H}_+^{n \times n}$, and $\mathbb{U}^{n \times n}$ denote the set of $n \times n$ Hermitian, positive semi-definite, and unitary matrices, respectively.
	$\mathbf{0}$ and $\mathbf{I}$ are the all-zero and identity matrices with appropriate size, respectively.
	% $\Re\{\cdot\}$ and $\mathrm{arg}(\cdot)$ take the real part and argument of a complex number, respectively.
	$\Re\{\cdot\}$ takes the real part of a complex number.
	$\tr(\cdot)$ and $\det(\cdot)$ evaluate the trace and determinant of a square matrix, respectively.
	$\diag(\cdot)$ constructs a square matrix by placing the inputs on the main (block) diagonal and zeros elsewhere.
	$\sv(\cdot)$ extracts the singular values of a matrix as a non-ascending vector.
	% $\card(\cdot)$ returns the cardinality of a set.
	$\conv(\cdot)$ returns the convex hull of arguments.
	$\lvert \cdot \rvert$, $\lVert \cdot \rVert$, and $\lVert \cdot \rVert _\mathrm{F}$ denote the absolute value, Euclidean norm, and Frobenius norm, respectively.
	$\sigma_n(\cdot)$ and $\lambda_n(\cdot)$ are the $n$-th largest singular value and eigenvalue, respectively.
	% $\boldsymbol{\sigma}(\cdot)$ and $\boldsymbol{\lambda}(\cdot)$ are the corresponding vectors.
	$(\cdot)^*$, $(\cdot)^\mathsf{T}$, $(\cdot)^\mathsf{H}$, $(\cdot)^\dagger$ $(\cdot)^{(r)}$, $(\cdot)^{\star}$ denote the conjugate, transpose, conjugate transpose (Hermitian), Moore-Penrose inverse, $r$-th iterated point, and stationary point, respectively.
	$[N]$ is a shortcut for $\{1,2,\ldots,N\}$.
	$(\cdot)_{[x:y]}$ is a shortcut for $(\cdot)_x,(\cdot)_{x+1},\ldots,(\cdot)_y$.
	$\odot$ denotes the element-wise (Hadamard) product.
	$\mathcal{CN}(\mathbf{0}, \mathbf{\Sigma})$ is the multivariate \gls{cscg} distribution with mean $\mathbf{0}$ and covariance $\mathbf{\Sigma}$.
	$\sim$ means ``distributed as''.
	\end{section}

	\begin{section}{System Model}
	Consider a \gls{bd}-\gls{ris} aided point-to-point \gls{mimo} system with $N_\mathrm{T}$ and $N_\mathrm{R}$ transmit and receive antennas, respectively, and $N_\mathrm{S}$ scattering elements at the \gls{bd}-\gls{ris}.
	This configuration is denoted as $N_\mathrm{T} \times N_\mathrm{S} \times N_\mathrm{R}$ throughout the paper.
	The \gls{bd}-\gls{ris} is modeled as an $N_\mathrm{S}$-port network \cite{Ivrlac2010} that divides into $G$ individual groups, where group $g \in [G]$ contains $N_g$ scattering elements interconnected by real-time reconfigurable components \cite{Shen2020a}.
	Apparently $N_\mathrm{S} = \sum_{g=1}^G N_g$.
	We assume all groups have the same size $N_g = L \triangleq N_\mathrm{S} / G, \ \forall g$ yet the discussion is extendable to \gls{bd}-\gls{ris} with unequal group size.
	To establish a theoretical performance bound, we consider a lossless asymmetric\footnote{A symmetric solution can be enforced by projection $\mathbf{\Theta} \gets (\mathbf{\Theta} + \mathbf{\Theta}^\mathsf{T})/2$ if necessary. We will compare the performance of symmetric and asymmetric \gls{bd}-\gls{ris} in Section~\ref{sc:simulation}.} \gls{bd}-\gls{ris} without mutual coupling as previously discussed in \cite{Li2023b,Li2023c,Bartoli2023}.
	The overall scattering matrix of the \gls{bd}-\gls{ris} is block-diagonal
	% \footnote{Following footnote \ref{fn:symmetric}, we do not constrain the scattering matrix to be symmetric in the paper. One can enforce such by $\mathbf{\Theta} \gets (\mathbf{\Theta} + \mathbf{\Theta}^\mathsf{T})/2$ if necessary.}
		\mathbf{\Theta} = \diag(\mathbf{\Theta}_1,\ldots,\mathbf{\Theta}_G),
	where $\mathbf{\Theta}_g \in \mathbb{U}^{L \times L}$ is the $g$-th unitary diagonal block modeling the response of group $g$.
	It is noteworthy that the conventional \gls{d}-\gls{ris} is an extreme case of \gls{bd}-\gls{ris} with group size $L=1$.
	% We highlight that conventional \gls{d}-\gls{ris} is an extreme case of \gls{bd}-\gls{ris} with group size $L=1$.
	Some potential physical architectures of \gls{bd}-\gls{ris} are illustrated in \cite[Fig. 3]{Shen2020a}, \cite[Fig. 5]{Li2023c}, and \cite[Fig. 2]{Nerini2024}, where the
	array geometry and
	circuit topology has been modeled in the scattering matrix.
	Let $\mathbf{H}_\mathrm{D} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{T}}$, $\mathbf{H}_\mathrm{B} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{S}}$, $\mathbf{H}_\mathrm{F} \in \mathbb{C}^{N_\mathrm{S} \times N_\mathrm{T}}$ denote the direct (i.e., transmitter-receiver), backward (i.e., \gls{ris}-receiver), and forward (i.e., transmitter-\gls{ris}) channels, respectively.
	The equivalent channel depends on the \gls{bd}-\gls{ris} scattering matrix
		\label{eq:channel}
		% \mathbf{H} = \mathbf{H}_\mathrm{D} + \mathbf{H}_\mathrm{B} \mathbf{\Theta} \mathbf{H}_\mathrm{F} = \mathbf{H}_\mathrm{D} + \sum_g \underbrace{\mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g}}_{\triangleq \mathbf{H}_g} = \mathbf{H}_\mathrm{D} + \sum_g \mathbf{H}_g,
		\mathbf{H} = \mathbf{H}_\mathrm{D} + \mathbf{H}_\mathrm{B} \mathbf{\Theta} \mathbf{H}_\mathrm{F} = \mathbf{H}_\mathrm{D} + \sum_g \mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g} \triangleq \mathbf{H}_\mathrm{D} + \sum_g \mathbf{H}_g,
	where $\mathbf{H}_{\mathrm{B},g} \in \mathbb{C}^{N_\mathrm{R} \times L}$ and $\mathbf{H}_{\mathrm{F},g} \in \mathbb{C}^{L \times N_\mathrm{T}}$ are the backward and forward channels associated with group $g$, corresponding to the $(g{-}1)L{+}1$ to $gL$ columns of $\mathbf{H}_\mathrm{B}$ and rows of $\mathbf{H}_\mathrm{F}$, respectively, and $\mathbf{H}_g \triangleq \mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g}$ is the indirect channel via group $g$.
	Since unitary matrices constitute an algebraic group with respect to multiplication, the scattering matrix of group $g$ can be decomposed as
		\mathbf{\Theta}_g = \mathbf{L}_g \mathbf{R}_g^\mathsf{H},
	where $\mathbf{L}_g, \mathbf{R}_g \in \mathbb{U}^{L \times L}$ are two unitary factor matrices.
	Let $\mathbf{H}_{\mathrm{B},g} = \mathbf{U}_{\mathrm{B},g} \mathbf{\Sigma}_{\mathrm{B},g} \mathbf{V}_{\mathrm{B},g}^\mathsf{H}$ and $\mathbf{H}_{\mathrm{F},g} = \mathbf{U}_{\mathrm{F},g} \mathbf{\Sigma}_{\mathrm{F},g} \mathbf{V}_{\mathrm{F},g}^\mathsf{H}$ be the \gls{svd} of the backward and forward channels, respectively.
	The equivalent channel can thus be rewritten as
		\label{eq:channel_svd}
		\mathbf{H} = \overbrace{\mathbf{H}_\mathrm{D} + \sum_g \mathbf{U}_{\mathrm{B},g} \mathbf{\Sigma}_{\mathrm{B},g} \underbrace{\mathbf{V}_{\mathrm{B},g}^\mathsf{H} \mathbf{L}_g \mathbf{R}_g^\mathsf{H} \mathbf{U}_{\mathrm{F},g}}_\text{backward-forward} \mathbf{\Sigma}_{\mathrm{F},g} \mathbf{V}_{\mathrm{F},g}^\mathsf{H}}^\text{direct-indirect}.

	% \begin{remark}
	We observe from \eqref{eq:channel_svd} that the \gls{bd}-\gls{ris} plays a crucial role in combining
	% needs to strike a balance between
	backward-forward (intra-group, multiplicative) channels and direct-indirect (inter-group, additive) channels.
	These two attributes are refined below and discussed in an example.
	% We conclude from \eqref{eq:channel_svd} that the scattering matrix of \gls{bd}-\gls{ris} provide two channel shaping potentials:
	% By analyzing \eqref{eq:channel_svd}, we conclude that the off-diagonal entries of the \gls{bd}-\gls{ris} scattering matrix provide two key potentials for \gls{mimo} channel shaping:
	% Channel rearrangement refers to rearrange and recombine the forward and backward channel branches by their strength.
	% Space alignment generalizes phase matching in \gls{siso} and \gls{miso} to the high-dimensional singular vector space in \gls{mimo}.
	\begin{itemize}
		\item \emph{Branch matching:} It refers to pairing and combining the branches (i.e., entries) of $\mathbf{H}_{\mathrm{B},g}$ and $\mathbf{H}_{\mathrm{F},g}$ through unitary transformation $\mathbf{\Theta}_g$.
			% That is,
			% % entries of $\mathbf{H}_{\mathrm{B},g}$ and $\mathbf{H}_{\mathrm{F},g}$ by their strength.
			% % rearranging and recombining the backward and forward channel branches (i.e., entries of $\mathbf{H}_\mathrm{B}$ and $\mathbf{H}_\mathrm{F}$)
			% % associated with each group
			% % by their strength.
			% % This unique feature of \gls{bd}-\gls{ris} exploits \textcolor{red}{spatial diversity} effectively thanks to in-group connections.
			% % This unique feature of \gls{bd}-\gls{ris} exploits the \textcolor{red}{spatial diversity} by rearranging and combining the backward and forward channel branches within each group.
			% In \gls{siso}\footnote{Following previous notation, here we denote the direct channel as $h_\mathrm{D}$, the backward and forward channels as $\mathbf{h}_\mathrm{B} \in \mathbb{C}^{N_\mathrm{S} \times 1}$ and $\mathbf{h}_\mathrm{F}^\mathsf{H} \in \mathbb{C}^{1 \times N_\mathrm{S}}$, whose $n$-th entries are $h_{\mathrm{B},n}$ and $h_{\mathrm{F},n}$, respectively.
			% The channels associated with \gls{bd}-\gls{ris} group $g$ are also denoted as $\mathbf{h}_{\mathrm{B},g} \in \mathbb{C}^{L \times 1}$ and $\mathbf{h}_{\mathrm{F},g}^\mathsf{H} \in \mathbb{C}^{1 \times L}$, whose $l$-th entries are $h_{\mathrm{B},g,l}$ and $h_{\mathrm{F},g,l}$, respectively.}, \gls{d}-\gls{ris} with perfect phase matching can provide a maximum indirect channel gain $g_\mathrm{D}=\sum_{n=1}^{N_\mathrm{S}} \lvert h_{\mathrm{B},n} \rvert \lvert h_{\mathrm{F},n} \rvert$,
			% % Thanks to in-group connections,
			% while \gls{bd}-\gls{ris} can exploit in-group connections and generalize it to
			% % this indirect channel gain to
			% $g_\mathrm{BD}=\sum_{g=1}^{G} \sum_{l=1}^{L} \lvert h_{\mathrm{B},g,\pi_{\mathrm{B},g}(l)} \rvert \lvert h_{\mathrm{F},g,\pi_{\mathrm{F},g}(l)} \rvert$,
			% % $\sum_{g=1}^{G} \sum_{l=1}^{L} \lvert h_{\mathrm{B},\pi_{\mathrm{B},g}(l)} \rvert \lvert h_{\mathrm{F},\pi_{\mathrm{F},g}(l)} \rvert$,
			% % $\sum_{g=1}^{G} \sum_{l=1}^{L} \lvert h_{\mathrm{B},g,l} \rvert \lvert h_{\mathrm{F},g,\pi_g(l)} \rvert$,
			% where $\pi_{\mathrm{B},g}$ and $\pi_{\mathrm{F},g}$ are arbitrary permutations of $[L]$.
			% Note the first summation is over groups and the second summation is over permuted channels.
			% By rearrangement inequality \cite[Inequality 10.2.1]{Hardy1952}, for \gls{bd}-\gls{ris} we have
			% 	g_\mathrm{BD} \le \sum_{g=1}^{G} \sum_{l=1}^{L} \lvert h_{\mathrm{B},g,\pi^\downarrow_{\mathrm{B},g}(l)} \rvert \lvert h_{\mathrm{F},g,\pi^\downarrow_{\mathrm{F},g}(l)} \rvert,
			% 	\label{eq:gain_rearrangement}
			% where $\pi^\downarrow_{\mathrm{B},g}$ and $\pi^\downarrow_{\mathrm{F},g}$ sort $\{\lvert h_{\mathrm{B},g,l} \rvert\}_{l=1}^L$ and $\{\lvert h_{\mathrm{F},g,l} \rvert\}_{l=1}^L$ in descending orders, respectively.
			% % are the descending permutations of $\lvert h_{\mathrm{B},g,l} \rvert$ and $\lvert h_{\mathrm{F},g,l} \rvert$, respectively.
			% That is to say, when the direct channel is blocked,
			% the maximum \gls{siso} channel gain is attained if the \gls{bd}-\gls{ris} pairs the $l$-th strongest backward and forward branches within each group.
			% % \footnote{Similarly, the minimum is attained when we pair the $l$-th strongest backward branch and the $L{-}l{+}1$-th strongest forward branch within each group.}
			% % We thus conclude that \gls{bd}-\gls{ris} exploits \textcolor{red}{spatial diversity} effectively thanks to in-group connections.
			% % By rearrangement inequality, the maximum channel gain is attained by pairing the $l$-th strongest backward and forward branches.

>>>>

% \title{Channel Shaping Using Beyond Diagonal RIS:\\Analysis, Optimization, and Enhanced Flexibility}
% \title{MIMO Channel Shaping and Rate Maximization via Beyond Diagonal RIS}
% \title{MIMO Channel Shaping Using Beyond Diagonal RIS: Singular Value Redistribution and Rate Maximization}
\title{Beyond Diagonal RIS: MIMO Channel Shaping and Rate Maximization}
\author{
	\IEEEauthorblockN{
		Yang Zhao, \IEEEmembership{Member, IEEE,}
		Hongyu Li, \IEEEmembership{Graduate Student Member, IEEE,}\\
		Massimo Franceschetti, \IEEEmembership{Fellow, IEEE,}
		and Bruno Clerckx, \IEEEmembership{Fellow, IEEE}
	}
	\thanks{
		Yang Zhao, Hongyu Li, and Bruno Clerckx are with the Department of Electrical and Electronic Engineering, Imperial College London, London SW7 2AZ, U.K. (e-mail: \{\href{mailto:yang.zhao18@imperial.ac.uk}{yang.zhao18}, \href{mailto:c.li21@imperial.ac.uk}{c.li21}, \href{mailto:b.clerckx@imperial.ac.uk}{b.clerckx}\}@imperial.ac.uk).

@@

	% geodesic \gls{rcg} algorithm is tailored for smooth optimization problems of asymmetric \gls{bd}-\gls{ris} with arbitrary group size, then invoked for the Pareto frontier of channel singular values.
	% To understand the gain from off-diagonal entries, we also derive analytical singular value bounds in rank-deficient and fully-connected scenarios.
	% As a side product, we tackle \gls{mimo} rate maximization problem by alternating between active beamformer (eigenmode transmission) and passive beamformer (\gls{rcg} algorithm) until convergence.
	% A low-complexity suboptimal solution based on channel shaping is also proposed, where the decoupled problem is formulated as channel power maximization and solved in closed form iteratively.
	% % We now show how channel shaping decouples the design and amounts to a low-complexity suboptimal solution.
	% % A low-complexity suboptimal solution decoupling both blocks is also proposed, where the channel shaping subproblem is formulated as channel power maximization and solved in closed form iteratively.
	% Theoretical analysis and numerical evaluation reveal that the shaping advantage of \gls{bd}-\gls{ris} increases with group size and \gls{mimo} dimensions, stemming from stronger channel rearrangement and space alignment capabilities.
\end{abstract}

@@

		% First, we quantify the capability of a \gls{bd}-\gls{ris} to reshape the \gls{mimo} point-to-point channel in terms of singular values.
		% The \emph{Pareto frontiers} are characterized by optimizing the {weighted sum of singular values}, where the weights can be positive, zero, or negative.
		% % This generalizes most singular value metrics and provides a powerful design framework.
		% The resulting singular value region generalizes most relevant metrics and provides an intuitive channel shaping benchmark.
		% We now discuss some analytical singular value bounds in rank-deficient and fully-connected scenarios, which help to demystify the gain from off-diagonal entries.
		% This is the first paper to answer the channel shaping question and highlight the \gls{bd}-\gls{ris} gain from a Pareto perspective.

@@

	\emph{Notation:}
	Italic, bold lower-case, and bold upper-case letters indicate scalars, vectors and matrices, respectively.
	$\jmath$ denotes the imaginary unit.
	% $\mathbb{C}$ represents the set of complex numbers.
	$\mathbb{H}^{n \times n}$, $\mathbb{H}_+^{n \times n}$, and $\mathbb{U}^{n \times n}$ denote the set of $n \times n$ Hermitian, positive semi-definite, and unitary matrices, respectively.
	% $\mathbf{0}$ and $\mathbf{I}$ are the all-zero and identity matrices with appropriate size, respectively.
	$\Re\{\cdot\}$ takes the real part of a complex number.
	% $\tr(\cdot)$ and $\det(\cdot)$ evaluate the trace and determinant of a square matrix, respectively.
	$\diag(\cdot)$ constructs a square matrix with arguments on the main (block) diagonal and zeros elsewhere.
	$\sv(\cdot)$, $\ran(\cdot)$, and $\ker(\cdot)$ evaluate the singular values (in descending order), column space (range), and kernel of a matrix, respectively.
	% $\sv(\cdot)$ extracts the singular values of a matrix as a non-ascending vector.
	% $\card(\cdot)$ returns the cardinality of a set.
	$\conv(\cdot)$ returns the convex hull of arguments.
	$\lvert \cdot \rvert$, $\lVert \cdot \rVert$, and $\lVert \cdot \rVert _\mathrm{F}$ denote the absolute value, Euclidean norm, and Frobenius norm, respectively.
	$\sigma_n(\cdot)$ and $\lambda_n(\cdot)$ are the $n$-th largest singular value and eigenvalue, respectively.
	% $\boldsymbol{\sigma}(\cdot)$ and $\boldsymbol{\lambda}(\cdot)$ are the corresponding vectors.
	$(\cdot)^*$, $(\cdot)^\mathsf{T}$, $(\cdot)^\mathsf{H}$, $(\cdot)^\dagger$ $(\cdot)^{(r)}$, $(\cdot)^{\star}$ denote the conjugate, transpose, conjugate transpose (Hermitian), Moore-Penrose inverse, $r$-th iterated point, and stationary point, respectively.
	$[N]$ is a shortcut for $\{1,2,\ldots,N\}$.
	$(\cdot)_{[x:y]}$ is a shortcut for $(\cdot)_x,(\cdot)_{x+1},\ldots,(\cdot)_y$.
	$\odot$ denotes the element-wise (Hadamard) product.
	$\mathcal{N}_{\mathbb{C}}(\mathbf{0}, \mathbf{\Sigma})$ is the multivariate \gls{cscg} distribution with mean $\mathbf{0}$ and covariance $\mathbf{\Sigma}$.
	$\sim$ means ``distributed as''.
\end{section}

\begin{section}{System Model}
	\begin{subsection}{BD-RIS}
		The \gls{bd}-\gls{ris} can be modeled as an $N_\mathrm{S}$-port network \cite{Ivrlac2010} that divides into $G$ individual groups, where group $g \in [G]$ contains $N_g$ scattering elements interconnected by real-time reconfigurable components \cite{Shen2020a}.
		Apparently $N_\mathrm{S} = \sum_{g=1}^G N_g$.
		Without loss of generality we assume equal size for all groups $N_g = L \triangleq N_\mathrm{S} / G, \ \forall g$ and no mutual coupling between elements.
		For asymmetric \gls{bd}-\gls{ris}, the overall scattering matrix is block-diagonal with unitary blocks\footnote{We assume asymmetric network by default to establish a benchmark for passive \gls{ris}. Some symmetric solutions (i.e., $\mathbf{\Theta} = \mathbf{\Theta}^\mathsf{T}$) will be discussed and their performance will be evaluated in Section~\ref{sc:simulation}.}
		\begin{equation}
			\label{eq:ris}
			\mathbf{\Theta} = \diag(\mathbf{\Theta}_1,\ldots,\mathbf{\Theta}_G),
		\end{equation}
		where $\mathbf{\Theta}_g \in \mathbb{U}^{L \times L}$ is the $g$-th diagonal block modeling the response of group $g$.
		It is noteworthy that \gls{d}-\gls{ris} is an extreme case of \eqref{eq:ris} with group size $L=1$, that is, each element is an individual.
		Some viable architectures of \gls{bd}-\gls{ris} are illustrated in \cite[Fig. 3]{Shen2020a}, \cite[Fig. 5]{Li2023c}, \cite[Fig. 2]{Nerini2024} where the array geometry and circuit topology are modeled in $\mathbf{\Theta}$.
	\end{subsection}

	\begin{subsection}{MIMO Point-to-Point Channel}
		Consider a \gls{bd}-\gls{ris} aided \gls{mimo} \gls{pc} with $N_\mathrm{T}$ and $N_\mathrm{R}$ transmit and receive antennas, respectively, and $N_\mathrm{S}$ scattering elements at the \gls{bd}-\gls{ris}.
		This configuration is denoted as $N_\mathrm{T} \times N_\mathrm{S} \times N_\mathrm{R}$.
		Let $\mathbf{H}_\mathrm{D} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{T}}$, $\mathbf{H}_\mathrm{B} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{S}}$, $\mathbf{H}_\mathrm{F} \in \mathbb{C}^{N_\mathrm{S} \times N_\mathrm{T}}$ denote the direct (i.e., transmitter-receiver), backward (i.e., \gls{ris}-receiver), and forward (i.e., transmitter-\gls{ris}) channels, respectively.
		The equivalent channel depends on the \gls{bd}-\gls{ris} scattering matrix
		\begin{equation}
			\label{eq:channel}
			\mathbf{H} = \mathbf{H}_\mathrm{D} + \mathbf{H}_\mathrm{B} \mathbf{\Theta} \mathbf{H}_\mathrm{F} = \mathbf{H}_\mathrm{D} + \sum_g \mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g} \triangleq \mathbf{H}_\mathrm{D} + \sum_g \mathbf{H}_g,
		\end{equation}
		where $\mathbf{H}_{\mathrm{B},g} \in \mathbb{C}^{N_\mathrm{R} \times L}$ and $\mathbf{H}_{\mathrm{F},g} \in \mathbb{C}^{L \times N_\mathrm{T}}$ are the backward and forward channels associated with group $g$, corresponding to the $(g{-}1)L{+}1$ to $gL$ columns of $\mathbf{H}_\mathrm{B}$ and rows of $\mathbf{H}_\mathrm{F}$, respectively, and $\mathbf{H}_g \triangleq \mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g}$ is the indirect channel via group $g$.
		Since unitary matrices constitute an algebraic group with respect to multiplication, the scattering matrix of group $g$ can be decomposed as
		\begin{equation}
			\mathbf{\Theta}_g = \mathbf{L}_g \mathbf{R}_g^\mathsf{H},
		\end{equation}
		where $\mathbf{L}_g, \mathbf{R}_g \in \mathbb{U}^{L \times L}$ are two unitary factor matrices.
		Let $\mathbf{H}_{\mathrm{B/F},g} = \mathbf{U}_{\mathrm{B/F},g} \mathbf{\Sigma}_{\mathrm{B/F},g} \mathbf{V}_{\mathrm{B/F},g}^\mathsf{H}$ be the \gls{svd} of the backward and forward channels, respectively.
		The equivalent channel is thus
		\begin{equation}
			\label{eq:channel_svd}
			\mathbf{H} = \overbrace{\mathbf{H}_\mathrm{D} + \sum_g \mathbf{U}_{\mathrm{B},g} \mathbf{\Sigma}_{\mathrm{B},g} \underbrace{\mathbf{V}_{\mathrm{B},g}^\mathsf{H} \mathbf{L}_g \mathbf{R}_g^\mathsf{H} \mathbf{U}_{\mathrm{F},g}}_\text{backward-forward} \mathbf{\Sigma}_{\mathrm{F},g} \mathbf{V}_{\mathrm{F},g}^\mathsf{H}}^\text{direct-indirect}.
		\end{equation}

		\begin{remark}
		In \eqref{eq:channel_svd}, the \gls{bd}-\gls{ris} performs a blockwise unitary transformation to combine the backward-forward (intra-group, multiplicative) channels and direct-indirect (inter-group, additive) channels.
		These two attributes are refined as:
		\begin{itemize}
			\item \emph{Branch matching:} It refers to pairing and combining the branches (i.e., entries) of $\mathbf{H}_{\mathrm{B},g}$ and $\mathbf{H}_{\mathrm{F},g}$ through $\mathbf{\Theta}_g$.
			\item \emph{Mode alignment:} It refers to aligning and ordering the modes (i.e., singular vectors) of $\{\mathbf{H}_g\}_{g \in [G]}$ with those of $\mathbf{H}_\mathrm{D}$ through $\{\mathbf{\Theta}_g\}_{g \in [G]}$.
		\end{itemize}
		\end{remark}

		\begin{example}[\gls{siso} channel gain maximization]
			\label{eg:siso}
			Denote the direct, backward, forward channels as $h_\mathrm{D}$, $\mathbf{h}_\mathrm{B} \in \mathbb{C}^{N_\mathrm{S} \times 1}$, and $\mathbf{h}_\mathrm{F}^\mathsf{H} \in \mathbb{C}^{1 \times N_\mathrm{S}}$, respectively.
			In this case, mode alignment boils down to phase matching and any $L \in [N_\mathrm{S}]$, including \gls{d}-\gls{ris}, suffices for perfect mode alignment using
			\begin{equation}
				\mathbf{\Theta}_{\textnormal{P-max},g}^\textnormal{SISO} = \frac{h_\mathrm{D}}{\lvert h_\mathrm{D} \rvert} \mathbf{V}_{\mathrm{B},g} \mathbf{U}_{\mathrm{F},g}^\mathsf{H}, \quad \forall g,
			\end{equation}
			where $\mathbf{V}_{\mathrm{B},g} = \bigl[\mathbf{h}_{\mathrm{B},g}/\lVert \mathbf{h}_{\mathrm{B},g} \rVert, \mathbf{N}_{\mathrm{B},g}\bigr] \in \mathbb{U}^{L \times L}$, $\mathbf{U}_{\mathrm{F},g} = \bigl[\mathbf{h}_{\mathrm{F},g}/\lVert \mathbf{h}_{\mathrm{F},g} \rVert, \mathbf{N}_{\mathrm{F},g}\bigr] \in \mathbb{U}^{L \times L}$, and $\mathbf{N}_{\mathrm{B/F},g} \in \mathbb{C}^{L \times (L-1)}$ are the orthonormal bases of kernels of $\mathbf{h}_{\mathrm{B/F},g}$.
			Evidently, the maximum channel gain is a function of $L$
			\begin{equation}
				\ \lvert h \rvert _{\max} = \lvert h_\mathrm{D} \rvert + \sum_g \sum_l \lvert h_{\mathrm{B},g,\pi_{\mathrm{B},g}(l)} \rvert \lvert h_{\mathrm{F},g,\pi_{\mathrm{F},g}(l)} \rvert,
			\end{equation}
			where $h_{\mathrm{B/F},g,l}$ are the $l$-th entries of $\mathbf{h}_{\mathrm{B/F},g}$, and $\pi_{\mathrm{B/F},g}$ are permutations of $[L]$ sorting their magnitude in similar orders.
			That is, the maximum \gls{siso} channel gain is attained when each \gls{bd}-\gls{ris} group, apart from phase shifting, matches the $l$-th strongest backward and forward channel branches therein.
			A larger $L$ provides more flexible branch matching and thus higher channel gain.
		\end{example}

		Example~\ref{eg:siso} clarifies the difference between branch matching and mode alignment and show their impacts on channel shaping.
		When it comes to \gls{mimo}, the advantage of \gls{bd}-\gls{ris} in branch matching improves since the number of available branches is proportional to $N_\mathrm{T} N_\mathrm{R}$. On the other hand, the limitation of \gls{d}-\gls{ris} in mode alignment worsens since each element can only apply a phase shift to the indirect channel of $N \triangleq \min(N_\mathrm{T}, N_\mathrm{R})$ modes.
	\end{subsection}

	\begin{subsection}{MIMO Interference Channel}
		We also consider a \gls{bd}-\gls{ris} aided \gls{mimo} \gls{ic} of $K$ transceiver pairs where each transmitter and receiver has $N_\mathrm{T}$ and $N_\mathrm{R}$ antennas, respectively, and the \gls{bd}-\gls{ris} has $N_\mathrm{S}$ scattering elements.
		This configuration is denoted as $(N_\mathrm{T} \times N_\mathrm{S} \times N_\mathrm{R})^K$.
		Let $\mathbf{H}_\mathrm{D}^{[kj]} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{T}}$, $\mathbf{H}_\mathrm{B}^{[k]} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{S}}$, $\mathbf{H}_\mathrm{F}^{[j]} \in \mathbb{C}^{N_\mathrm{S} \times N_\mathrm{T}}$ denote the direct channel from transmitter $j$ to receiver $k$, the backward channel of receiver $k$, and the forward channel of transmitter $j$, respectively, where $(j,k) \in [K]^2$.
		Assume all transmitter-\gls{ris}-receiver paths share the same \gls{bd}-\gls{ris} scattering matrix $\mathbf{\Theta}$.
		The equivalent channel from transmitter $j$ to receiver $k$ is
		\begin{equation}
			\mathbf{H}^{[kj]} = \mathbf{H}_\mathrm{D}^{[kj]} + \mathbf{H}_\mathrm{B}^{[k]} \mathbf{\Theta} \mathbf{H}_\mathrm{F}^{[j]} = \mathbf{H}_\mathrm{D}^{[kj]} + \sum_g \mathbf{H}_{\mathrm{B},g}^{[k]} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g}^{[j]},
			\label{eq:channel_equivalent_ic}
		\end{equation}
		where $\mathbf{H}_{\mathrm{B},g}^{[k]} \in \mathbb{C}^{N_\mathrm{R} \times L}$ and $\mathbf{H}_{\mathrm{F},g}^{[j]} \in \mathbb{C}^{L \times N_\mathrm{T}}$ are associated with \gls{ris} group $g$, corresponding to the $(g{-}1)L{+}1$ to $gL$ columns of $\mathbf{H}_{\mathrm{B},g}^{[k]}$ and rows of $\mathbf{H}_{\mathrm{F},g}^{[j]}$, respectively,
	\end{subsection}
\end{section}

\begin{section}{Channel Shaping}
	In this section, we first provide an example demonstrating the \gls{mimo} channel shaping advantages of \gls{bd}-\gls{ris} over \gls{d}-\gls{ris}, then derive some analytical bounds related to channel singular values under specific channel conditions, and provide closed-form \gls{bd}-\gls{ris} solution for several cases of interest.
	Finally, we propose a numerical method to optimize the \gls{bd}-\gls{ris} for a special class of singular value function under general channel conditions.
